---
title: "Code used for analysis"
author: |
    | Chloe You 
    | xinyuan.you@stat.ubc.ca
date: "`r format(Sys.time(), '%B %Y')`"
output: pdf_document
---

#Simulation: Normal, Student-t, and Cauchy Prior Comparisons

```{r, message=F, warning=F}
library(MCMCvis)
library(rjags)
set.seed(1)
```

## Generate Data

Let $x_i$ be the indicator variable of being in the treatment group or not. $p_{control}$ and $p_{trtment}$ is the true probability of the subject being in each of the group. We simulate data via the following process:

$$
x_{i} \sim Bernoulli(0.5) 
$$
$$
y_i \sim Binomial(n, p_i) 
$$
$$
p_i = \begin{cases}
 p_{control} & x_i=0 \\
 p_{trtment}=p_{control} RR_{true}   & x_i=1
 \end{cases} 
$$
We set sample size $n=100$, $p_{control}=0.3$, and $RR_{true}=0.7$. We set $\beta_0 = -2$ for easier computation. 


```{r}
sim = list()

init = 1
for(r in c(0.7)){
  for (n in c(100)){
    sim[[init]] = data.frame(x=numeric(n),p=numeric(n),y=numeric(n))
    for(i in c(1:n)){
      sim[[init]]$x[i] = rbinom(1,1,prob=0.5)
      sim[[init]]$p[i] = (1-sim[[init]]$x[i])*0.3 + sim[[init]]$x[i]*0.3*r
      sim[[init]]$y[i] = rbinom(1, 1, prob=sim[[init]]$p[i])
    }
    init = init + 1
  }
}
```


## Prior Specification and Statistical Model

### Prior 1: $\beta_1 \sim$ Normal(0, $0.35^2$)

```{r}
genmod.string <- "model{
  
  ### prior distribution
  b1 ~ dnorm(0, 8.163)

  ### target parameter
  RR <- exp(b1)

  ### statistical model
  for (i in 1:n) {
    y[i] ~ dbern(p.bound[i])
    p.bound[i] <- max(0, min(1, p[i]))
    log(p[i]) <- -2 + b1*x[i]
  }  
}"

### generative model, data go in

mod <- jags.model(textConnection(genmod.string),
                  data=list(x= sim[[1]]$x, y= sim[[1]]$y,n= nrow(sim[[1]])),
                  n.chains=3)
# n.chains = the number of parallel chains for the model

###  MC output comes out

opt.JAGS <- coda.samples(mod, 
                 variable.names=c("RR"), 
                 n.iter=10000)

MCMCsummary(opt.JAGS)
```

### Prior 2: $\beta_1 \sim$ Cauchy(0, $2.5^2$)

```{r}
genmod.string <- "model{
  
  ### prior distribution
  b1 ~ dt(0, 0.16, 1)

  ### target parameter
  RR <- exp(b1)

  ### statistical model
  for (i in 1:n) {
    y[i] ~ dbern(p.bound[i])
    p.bound[i] <- max(0, min(1, p[i]))
    log(p[i]) <- -2 + b1*x[i]
  }  
}"

### generative model, data go in

mod <- jags.model(textConnection(genmod.string),
                  data=list(x= sim[[1]]$x, y= sim[[1]]$y,n= nrow(sim[[1]])),
                  n.chains=3)
# n.chains = the number of parallel chains for the model

###  MC output comes out

opt.JAGS <- coda.samples(mod, 
                 variable.names=c("RR"), 
                 n.iter=10000)

MCMCsummary(opt.JAGS)
```

### Prior 3: $\beta_1 \sim$ $t_7(0, 0.35^2)$
```{r}
genmod.string <- "model{
  
  ### prior distribution
  b1 ~ dt(0, 8.16, 7)

  ### target parameter
  RR <- exp(b1)

  ### statistical model
  for (i in 1:n) {
    y[i] ~ dbern(p.bound[i])
    p.bound[i] <- max(0, min(1, p[i]))
    log(p[i]) <- -2 + b1*x[i]
  }  
}"

### generative model, data go in

mod <- jags.model(textConnection(genmod.string),
                  data=list(x= sim[[1]]$x, y= sim[[1]]$y,n= nrow(sim[[1]])),
                  n.chains=3)
# n.chains = the number of parallel chains for the model

###  MC output comes out

opt.JAGS <- coda.samples(mod, 
                 variable.names=c("RR"), 
                 n.iter=10000)

MCMCsummary(opt.JAGS)
```

### Non-Bayesian Estimates
```{r}
nonbay= glm(formula = sim[[1]]$y ~ sim[[1]]$x, family = binomial(link = log))
summary(nonbay)
exp(confint(nonbay))
```



